Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-06-05 neurips-20/cvx-flats.tex
pass through Sec 4

The \emph{level set} of $\Gamma$ at value $r\in\R$ is $\Gamma_r := \{p \in \simplex : r \in \Gamma(p)\}$.
We call a property $\Gamma: \simplex \toto \R$ \emph{discrete} if $\R$ is a finite set.
A property is \emph{single-valued} if $|\Gamma(p)| = 1$ for all $p\in\simplex$.%, in which case we write $\Gamma\to\R$.\botodo{We do?}\jessiet{Commenting out.  We only used once in the appendix, do I changed that instance.}
We use $\Gamma := \prop{L}$ to denote the unique property elicited by $L$ from eq.~\eqref{eq:elic}.\footnote{Note that $\prop{L}$ is well-defined if and only if $L \in \L$.}
Finally, when multiple properties are in play, we typically denote the target property by $\gamma$, and the surrogate by $\Gamma$.

@@

\btw{cut this next paragraph. but still want to answer the question of ``how does elicitation relate to calibration?''}
%As we show next, calibration implies indirect elicitation, which means the convex elicitation complexity always lower bounds the convex calibration dimension in the setting of discrete target losses.
%Yet despite this fact, the state-of-the-art lower bound for convex calibration dimension, the \emph{feasible subspace dimension} of \citet{ramaswamy2016convex}, is a special case of our lower bound for convex elicitation complexity (Theorem~\ref{thm:cvx-flats}), meaning we achieve the same or even higher lower bounds.
%(See the examples in \S~\ref{subsec:examples-finite}.)
%Moreover, convex elicitation complexity also applies beyond the discrete target loss setting, as we demonstrate in \S~\ref{sec:contin-consis}.

In related work,
\cite[Corollary 10]{agarwal2015consistent} provides a necessary condition for the direct convex elicitation of single-valued properties, whose yielding bounds on the dimensionality of level sets.
Moreover, \citet{finocchiaro2019embedding} study surrogate losses which \emph{embed} a discrete loss in some sense, which is a special case of indirect elicitation.
They construct a calibrated link functions from embeddings, which together with observations about indirect elicitation, imply that for polyhedral (piecewise linear and convex) losses, calibration and indirect elicitation are equivalent.
\citet{finocchiaro2020embedding} further introduce the notion of \emph{embedding dimension}, which is a lower bound on both convex elicitation complexity of discrete properties and convex calibration dimension of discrete losses.
\btw{Bo: commented out what I felt was an unnecessary sentence}
%Since embedding is a special case of indirect elicitation, we have the chain of inequalities, embedding dimension $\leq$  ``polyhedral calibration dimension'' $\leq$ convex calibration dimension; it is an open problem whether any of these inequalities can be strict.

@@

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
  \btw{Note: don't actually need $\gamma$ nonempty here.}
\end{lemma}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \inf_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

We now turn to the question of bounding the prediction dimension of a consistent convex surrogate.
From Theorem \ref{thm:consistent-implies-indir-elic}, given a target property $\gamma$ or loss $\ell$ with $\gamma = \prop{\ell}$, this task reduces to lower bounding the prediction dimension of a convex surrogate indirectly eliciting $\gamma$.
We now explore a tool, Theorem \ref{thm:cvx-flats}, for proving such convex-elicitation lower bounds.
The key idea, crystallized from the proof of \citet[Theorem 16]{ramaswamy2016convex}, is to consider a particular distribution~$p$ and surrogate prediction $u \in \reals^d$ with is optimal for $p$.
Theorem \ref{thm:cvx-flats} will show that if $d$ is small, then the level set $\{p : u \in \argmin_{u'} \exploss{L}{u'}{p}\}$ must be large; in fact, it must roughly contain a high-dimensional affine subspace, or \emph{flat}.
But by definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The point of this result is the contrapositive: if $\gamma$ has a level set intricate enough to not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.

@@

We now state our lower bound.

@@

The proof takes its insipration from~\citet[Theorem 16]{ramaswamy2016convex}, and extracts a witness $W$ of the optimality of a surrogate prediction $u$.
This witness $W$ takes the form of a matrix whose columns are selection of subgradients of $L$, and thus $u$ must also be optimal for any other distribution in the flat defined by the kernel of $W$.
As we show next, the original statement of~\citet[Theorem 16]{ramaswamy2016convex}, which uses the stronger notion of calibration and thus applies only to discrete target losses, actually follows as a special case of Theorem~\ref{thm:cvx-flats}. \jessiet{How to rephrase this?}

@@

\begin{lemma}[\citep{frongillo2018elicitation}]
	\label{lem:elic-complex-bayes-concave}
	Suppose the loss $L$ elicits a single-valued property $\Gamma:\simplex\toto\R$.
	Let $\lbar$ be the Bayes risk of $L$.
	Then for any $p,p'\in\simplex$ with $\Gamma(p)\neq\Gamma(p')$, we have $\lbar(\lambda p + (1-\lambda) p') > \lambda \lbar(p) + (1-\lambda) \lbar(p')$ for all $\lambda\in(0,1)$.
\end{lemma}

>>>>

The \emph{level set} of $\Gamma$ at value $r\in\R$ is $\Gamma_r := \{p \in \simplex : r \in \Gamma(p)\}$.
We call a property $\Gamma: \simplex \toto \R$ \emph{discrete} if $\R$ is a finite set.
A property is \emph{single-valued} if $|\Gamma(p)| = 1$ for all $p\in\simplex$, in which case we write $\Gamma\to\R$.\botodo{We do?}
We use $\Gamma := \prop{L}$ to denote the unique property elicited by $L$ from eq.~\eqref{eq:elic}.\footnote{\bo{added} Note that $\prop{L}$ is well-defined if and only if $L \in \L$.}
Finally, when multiple properties are in play, we typically denote the target property by $\gamma$, and the surrogate by $\Gamma$.

@@

\btw{cut this? but still want to answer the question of ``how does elicitation relate to calibration?''}
%As we show next, calibration implies indirect elicitation, which means the convex elicitation complexity always lower bounds the convex calibration dimension in the setting of discrete target losses.
%Yet despite this fact, the state-of-the-art lower bound for convex calibration dimension, the \emph{feasible subspace dimension} of \citet{ramaswamy2016convex}, is a special case of our lower bound for convex elicitation complexity (Theorem~\ref{thm:cvx-flats}), meaning we achieve the same or even higher lower bounds.
%(See the examples in \S~\ref{subsec:examples-finite}.)
%Moreover, convex elicitation complexity also applies beyond the discrete target loss setting, as we demonstrate in \S~\ref{sec:contin-consis}.

In related work,
\citet{finocchiaro2019embedding} study surrogate losses which \emph{embed} a discrete loss in some sense, which is a special case of indirect elicitation.
They construct a calibrated link functions from embeddings, which together with observations about indirect elicitation, imply that for polyhedral (piecewise linear and convex) losses, calibration and indirect elicitation are equivalent.
\citet{finocchiaro2020embedding} further introduce the notion of \emph{embedding dimension}, which is a lower bound on both convex elicitation complexity of discrete properties and convex calibration dimension of discrete losses.
\btw{Bo: commented out what I felt was an unnecessary sentence}
%Since embedding is a special case of indirect elicitation, we have the chain of inequalities, embedding dimension $\leq$  ``polyhedral calibration dimension'' $\leq$ convex calibration dimension; it is an open problem whether any of these inequalities can be strict.

@@

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the regret with respect to $\ell$, the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is consistent with respect to $\ell$ if and only if $(L,\psi)$ is $\propdis$-consistent with respect to $\gamma := \prop{\ell}$.
  \btw{Note: don't actually $\gamma$ nonempty here.}
\end{lemma}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \inf_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

We now turn to the question of bounding the prediction dimension of a consistent convex surrogate.
From Theorem \ref{thm:consistent-implies-indir-elic}, given a target property $\gamma$ or loss $\ell$ with $\gamma = \prop{\ell}$, this task reduces to lower bounding the prediction dimension of a convex surrogate indirectly eliciting $\gamma$.
We now explore a tool, Theorem \ref{thm:cvx-flats}, for proving such convex-elicitation lower bounds.
The key idea, crystallized from the proof of \citet[Theorem 16]{ramaswamy2016convex} (a similar technique appears in \citet[Theorem 9]{agarwal2015consistent}), is to consider a particular distribution~$p$ and surrogate prediction $u \in \reals^d$ with is optimal for $p$.
Theorem \ref{thm:cvx-flats} will show that if $d$ is small, then the level set $\{p : u \in \argmin_{u'} \exploss{L}{u'}{p}\}$ must be large; in fact, it must roughly contain a high-dimensional affine subspace, or \emph{flat}.
But by definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The point of this result is the contrapositive: if $\gamma$ has a level set intricate enough to not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.

@@

We now state our elicitation lower bound, Theorem \ref{thm:cvx-flats}.
By combining with Theorem \ref{thm:consistent-implies-indir-elic}, it will be used to imply consistency bounds.
\botodo{Added comparison to AA15; please proofread!}
A result of a similar flavor is \citet[Theorem 9]{agarwal2015consistent}, which lower-bounds the affine dimension of level sets of $\prop{L}$.
Theorem \ref{thm:cvx-flats} gives a more powerful criterion, namely containment of flats of certain dimension in the level sets, and this is needed to give good bounds for some of our examples.
Theorem \ref{thm:cvx-flats} also gives a lower bound for any given target $\ell$ or $\gamma$, rather than a given surrogate $L$ (although an analogous implication of their Theorem 9 may have been known to \citet{agarwal2015consistent}).
We also note that the lower bound on elicitation complexity in \citet{agarwal2015consistent} is not translated into a lower bound for \emph{consistent} surrogates, which is the main contribution of this work.

@@

The proof takes its insipration from~\citet[Theorem 16]{ramaswamy2016convex}, and extracts a witness $W$ of the optimality of a surrogate prediction $u$.
This witness $W$ takes the form of a matrix whose columns are selection of subgradients of $L$, and thus $u$ must also be optimal for any other distribution in the flat defined by the kernel of $W$.
As we show next, the original statement of~\citet[Theorem 16]{ramaswamy2016convex}, which uses the stronger notion of calibration and thus applies only to discrete target losses, actually follows as a special case of Theorem~\ref{thm:cvx-flats}. \jessiet{How to rephrase this?}
\botodo{I suggest cutting this paragraph entirely if we keep the full proof in. We mention above that it takes inspiration from Ramaswamy.}

@@

\begin{lemma}[\citep{frongillo2018elicitation}]
	\label{lem:elic-complex-bayes-concave}
	Suppose the loss $L$ elicits $\Gamma:\simplex\to\R$.
	Let $\lbar$ be the Bayes risk of $L$.
	Then for any $p,p'\in\simplex$ with $\Gamma(p)\neq\Gamma(p')$, we have $\lbar(\lambda p + (1-\lambda) p') > \lambda \lbar(p) + (1-\lambda) \lbar(p')$ for all $\lambda\in(0,1)$.
\end{lemma}
